\bf The \texttt{forestecology} R package for modeling interspecies competition between trees

### 2021-02-16 etc/paper/forestecology.tex
Done "Identify all focal and corresponding competitor trees" section

@@

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Define buffer region trees.
  \item
    Add spatial cross-validation block information.
  \end{enumerate}
\item
  Compute focal versus competitor tree information.
\item
  Fit model and make predictions.
\item
  Visualize model results.
\item
  Additionally: Evaluate model performance using spatial
  cross-validation.
\item
  Additionally: Evaluate the effect of competitor species identity using
  permutation tests.
\end{enumerate}

@@

The first step in the our analysis sequence is to compute the growth of
trees using data from two censuses. The \texttt{compute\_growth()}
function computes growth assuming census data that follows ForestGEO
standards. Despite such standards, minor variations will still exist
between sites thereby necessitating some data wrangling and checking.
For example, the SCBI site records all dbh's in millimeters, whereas the
Michigan Big Woods site records them in centimeters
\citet{andersonteixeira_ctfs-forestgeo_2015}
\citet{allen_michigan_2020}. The data format of other sites may be such
that our \texttt{compute\_growth()} function doesn't work at all.
However, in the end all that matters is that the growth of all trees is
saved in a data frame of class \texttt{sf} whereby the geolocation of
each tree is presented in a \texttt{geometry} variable of type
\texttt{\textless{}POINT\textgreater{}} and at a minimum the data
contains the following variables: a variable uniquely identifying each
tree-stem, \texttt{sp} of type \texttt{fct} factor identifying species,
\texttt{dbh1} and \texttt{dbh2} of type \texttt{dbl} quantifying the dbh
at earlier and later census, and \texttt{growth} of type \texttt{dbl}
double quantifying the average annual growth in centimeters.

We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
existed on GitHub on November 20, 2020. After selecting only relevant
variables, we perform a few additional data wrangling steps: convert the
variable with the date of measurement to be of type \texttt{date},
convert dbh to be in centimeters\footnote{A rule of thumb to determine
  the units of dbh is check if the smallest non-zero and non-missing
  measurement is 1 or 10. If the former, then centimeters. If the later,
  then millimeters. This is because ForestGEO protocols state that only
  trees with dbh greater or equal to 1cm should be included in censuses.},
convert the \texttt{sp} variable containing species information from
type \texttt{chr} character to \texttt{fct} factor (we will discuss the
need for this in Section \ref{spatial-cross-validation}). Furthermore,
in order to speed up computation for purposes of this examplt, we only
consider a 9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx}
from 0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
0--640.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi}
\CommentTok{## Simple feature collection with 7954 features and 8 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 299.9 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 9}
\CommentTok{##    stemID sp     dbh1 codes1 status  dbh2 codes2  growth     geometry}
\CommentTok{##     <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>    <dbl>      <POINT>}
\CommentTok{##  1      4 nysy  13.6  M      A       14.2 M       0.103  (14.2 428.5)}
\CommentTok{##  2      5 havi   8.8  M      A        9.6 M;P     0.150   (9.4 436.4)}
\CommentTok{##  3      6 havi   3.25 NULL   A        4   M       0.140     (1.3 434)}
\CommentTok{##  4     77 qual  65.2  M      A       66   M       0.141  (34.7 307.2)}
\CommentTok{##  5     79 tiam  47.7  M      A       46.8 M      -0.161    (40 381.1)}
\CommentTok{##  6     80 caca   5.15 M      A        6.5 M       0.253  (38.7 421.7)}
\CommentTok{##  7     96 libe   2.3  J;M    A        3.7 M       0.262      (60 310)}
\CommentTok{##  8    100 caca   5.09 NULL   A       NA   DN     NA      (52.5 476.3)}
\CommentTok{##  9    101 litu  65.4  M      A       68.4 M       0.552  (47.1 567.3)}
\CommentTok{## 10    102 astr   1.99 NULL   A        2.5 M       0.0954 (40.8 575.5)}
\CommentTok{## # ... with 7,944 more rows}
\end{Highlighting}
\end{Shaded}

The output \texttt{growth\_scbi} is a single data frame of class
\texttt{sf} that includes a numerical \texttt{growth} reflecting the
average annual growth in dbh (in cm) for all trees that were alive at
both time points as well a \texttt{geometry} variable encoding each
tree's geolocation. Furthermore, variables that (in theory) remain
unchanged between censuses appear only once, such as location variables
\texttt{gx} and \texttt{gy}; as well as species-related variables.
Variables that should change between censuses are suffixed with
\texttt{1} and \texttt{2} indicating the earlier and later censuses,
such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.

@@

\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
  \KeywordTok{sf_polygon}\NormalTok{(fold1),}
  \KeywordTok{sf_polygon}\NormalTok{(fold2)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{define-focal-versus-competitor-trees}{%
\subsection{Define focal versus competitor
trees}\label{define-focal-versus-competitor-trees}}

@@

Next we define \texttt{focal\_vs\_comp} data frames which connects each
focal tree in the \texttt{growth\_df} data frames to the trees in its
competitive neighborhood range as defined by the \texttt{comp\_dist}
constant. So for example, if \texttt{growth\_df} consisted of two focal
trees with two and three neighbors with \texttt{comp\_dist}
respectively, \texttt{focal\_vs\_comp} would be a data frame of 5 rows
connecting each focal tree to it's competitors. The
\texttt{create\_focal\_vs\_comp()} function makes this connection taking
as inputs the \texttt{growth\_df} data frame; the \texttt{comp\_dist}
constant defining competitive range; \texttt{cv\_grid\_sf}, giving the
cross validation grid; and the \texttt{id} variable.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

TODO: Figure out how to show this data frame's contents.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

TODO: Figure out how to show this data frame's contents.

Next we fit the following linear model to the dbh of each focal tree.
Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
species groups. We modeled the growth in diameter per year \(y_{ij}\)
(in centimeters per year) of the \(i^{th}\) tree of focal species group
\(j\) as a linear model \(f\) of the following covariates
\(\vec{x}_{ij}\)

@@

Using these posterior means, we then use the posterior predictive
distribution to obtain fitted/predicted values \(\widehat{y}\) of the
dbh for each focal tree using the \texttt{predict\_bayesian\_model()}.
These \(\widehat{y}\) can then be compared to the observed \(y\) dbh's
to compute the root mean-square error, a measure of a model's predictive
error which has the same units as the observed data \(y\).

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.148145}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1505383}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1281398}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=1\linewidth]{Figures/scbi-model-residuals-1}

\caption{Predicted versus observed growth.}\label{fig:scbi-model-residuals}
\end{figure}

{\centering \includegraphics[width=0.8\linewidth]{Figures/scbi-model-residuals-2-1}

\caption{Spatial distribution of residuals for model applied to SCBI data.}\label{fig:scbi-model-residuals-2}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta0-1}

\caption{Posterior distribution of beta0.}\label{fig:bw-posterior-viz-beta0}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta-dbh-1}

\caption{Posterior distribution of betadbh.}\label{fig:bw-posterior-viz-beta-dbh}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-lambda-1}

\caption{Posterior distribution of lambda's for Big Woods.}\label{fig:bw-posterior-viz-lambda}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{Figures/scbi-posterior-viz-lambda-1}

\caption{Posterior distribution of lambda's for SCBI.}\label{fig:scbi-posterior-viz-lambda}
\end{figure}

@@

\hypertarget{big-woods-3}{%
\subsubsection{Big Woods}\label{big-woods-3}}

@@

\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1532316}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-3}{%
\subsubsection{SCBI}\label{scbi-3}}

@@

\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.144608}
\end{Highlighting}
\end{Shaded}

>>>>

\hypertarget{model}{%
\subsection{Model}\label{model}}

Describe model specifics.

@@

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Define buffer region trees.
  \item
    Add spatial cross-validation block information.
  \end{enumerate}
\item
  Identify all focal and corresponding competitor trees.
\item
  Fit model and make predictions.
\item
  Additionally: Evaluate model performance using spatial
  cross-validation.
\item
  Additionally: Evaluate the effect of competitor species identity using
  permutation tests.
\end{enumerate}

@@

The first step in the our analysis sequence is to compute the growth of
trees using data from two censuses. The \texttt{compute\_growth()}
function computes growth assuming census data that follows ForestGEO
standards. Despite such standards, minor variations will still exist
between sites thereby necessitating some data wrangling and checking.
For example, the SCBI site records all DBH's in millimeters, whereas the
Michigan Big Woods site records them in centimeters
\citet{andersonteixeira_ctfs-forestgeo_2015}
\citet{allen_michigan_2020}. The data format of other sites may be such
that our \texttt{compute\_growth()} function doesn't work at all.
However, in the end all that matters is that the growth of all trees is
saved in a data frame of class \texttt{sf} whereby the geolocation of
each tree is presented in a \texttt{geometry} variable of type
\texttt{\textless{}POINT\textgreater{}} and at a minimum the data
contains the following variables: a variable uniquely identifying each
tree-stem, \texttt{sp} of type \texttt{fct} factor identifying species,
\texttt{dbh1} and \texttt{dbh2} of type \texttt{dbl} quantifying the DBH
at earlier and later census, and \texttt{growth} of type \texttt{dbl}
double quantifying the average annual growth in centimeters.

We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
existed on GitHub on November 20, 2020. After selecting only relevant
variables, we perform a few additional data wrangling steps: convert the
variable with the date of measurement to be of type \texttt{date},
convert DBH to be in centimeters\footnote{A rule of thumb to determine
  the units of DBH is check if the smallest non-zero and non-missing
  measurement is 1 or 10. If the former, then centimeters. If the later,
  then millimeters. This is because ForestGEO protocols state that only
  trees with DBH greater or equal to 1cm should be included in censuses.},
convert the \texttt{sp} variable containing species information from
type \texttt{chr} character to \texttt{fct} factor (we will discuss the
need for this in Section \ref{spatial-cross-validation}). Furthermore,
in order to speed up computation for purposes of this example, we only
consider a 9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx}
from 0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
0--640.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi}
\CommentTok{## Simple feature collection with 7954 features and 8 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 299.9 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 9}
\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth     geometry}
\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>      <POINT>}
\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103 (14.2 428.5)}
\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150  (9.4 436.4)}
\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140    (1.3 434)}
\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141 (34.7 307.2)}
\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161   (40 381.1)}
\CommentTok{## # ... with 7,949 more rows}
\end{Highlighting}
\end{Shaded}

The output \texttt{growth\_scbi} is a single data frame of class
\texttt{sf} that includes a numerical \texttt{growth} reflecting the
average annual growth in DBH (in cm) for all trees that were alive at
both time points as well a \texttt{geometry} variable encoding each
tree's geolocation. Furthermore, variables that (in theory) remain
unchanged between censuses appear only once, such as location variables
\texttt{gx} and \texttt{gy}; as well as species-related variables.
Variables that should change between censuses are suffixed with
\texttt{1} and \texttt{2} indicating the earlier and later censuses,
such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.

@@

\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }
               \DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{identify-all-focal-and-corresponding-competitor-trees}{%
\subsection{Identify all focal and corresponding competitor
trees}\label{identify-all-focal-and-corresponding-competitor-trees}}

The next step in our analysis sequence is to identify all focal trees
and their corresponding competitor trees. The
\texttt{create\_focal\_vs\_comp()} functions performs these tasks and
returns a new data frame of type \texttt{sf} containing this
information. On top of the previously discussed arguments
\texttt{comp\_dist} defining the competition neighborhood and
\texttt{id} indicating which variable in the data frame uniquely
identifies each tree-stem, this function also requires an \texttt{sf}
object representation of the spatial cross-validation blocks/folds; in
our case, this was manually encoded in the \texttt{blocks\_scbi} in
Section \ref{spatial-information} while in our Appendix we present an
example where this was performed using \texttt{spatialBlock()} from the
\texttt{blockCV} package. We present the resulting data frame below with
the \texttt{foldID} variable omitted for compactness of presentation.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{foldID)}
\CommentTok{## # A tibble: 6,296 x 6}
\CommentTok{##   focal_ID focal_sp   dbh     geometry growth comp             }
\CommentTok{##      <dbl> <fct>    <dbl>      <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     13.6  (14.2 428.5)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      8.8   (9.4 436.4)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam     47.7    (40 381.1) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca      5.15 (38.7 421.7)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe      2.3      (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows}
\end{Highlighting}
\end{Shaded}

TODO: Below reconcile the number of rows as they off by one from
\texttt{growth\_scbi} pipe \texttt{filter(!is.na(growth)\ \&\ !buffer)}.
perhaps by removing NA's in the \texttt{growth\_scbi} stage.

The resulting data frame \texttt{focal\_vs\_comp\_scbi} has 6296 rows,
representing the subset of the 7954 trees in \texttt{growth\_scbi} that
will be considered as focal trees and thus have their growths modeled.
Recall from Section \ref{spatial-information} this consists all trees
that are not part of the buffer region in Figure
\ref{fig:scbi-spatial-information}. Two new variables \texttt{focal\_ID}
and \texttt{focal\_sp} related to tree-stem identification and species
information. Most notably however is a new variable \texttt{comp} which
contains information on all competitor trees for a given focal tree
saved in list-column format, a feature of the \texttt{tidyr} package
\citet{tidyr_package}. For example, we drill-down on the tree with
\texttt{focal\_ID} equal to 4. It has 20 competitor trees each described
by 4 variables as indicated by the fact that \texttt{comp} is a
\texttt{\textless{}tibble\ {[}20\ ×\ 4{]}\textgreater{}}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp)}
\CommentTok{## # A tibble: 1 x 3}
\CommentTok{##   focal_ID   dbh comp             }
\CommentTok{##      <dbl> <dbl> <list>           }
\CommentTok{## 1        4  13.6 <tibble [20 x 4]>}
\end{Highlighting}
\end{Shaded}

Using the \texttt{unnest()} function from the \texttt{tidyr} package, we
can flatten list-column into regular columns. We observe that for the
same focal tree with DBH equal to 13.65cm, we have information on all 20
competitor trees whose \texttt{dist} distance to the focal tree is less
than or equal to 7.5, including their unique tree-stem ID number, their
species, and their basal area (in m\(^2\)) calculated as
\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where \(DBH\) is the value
from the earlier census in cm. Saving our focal versus competitor
information in list-column minimizes redundancy since we do not repeat
information on the focal tree 20 times.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 20 x 6
##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area
##      <dbl> <dbl>   <dbl> <dbl> <fct>             <dbl>
## 1        4  13.6    1836  7.48 tiam            0.0176 
## 2        4  13.6    1847  2.81 nysy            0.00332
## 3        4  13.6    1848  1.62 nysy            0.00396
## 4        4  13.6    1849  2.62 nysy            0.00535
## 5        4  13.6    1850  2.98 havi            0.00472
## # ... with 15 more rows
\end{verbatim}

@@

Next we fit the following linear model to the DBH of each focal tree.
Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
species groups. We modeled the growth in diameter per year \(y_{ij}\)
(in centimeters per year) of the \(i^{th}\) tree of focal species group
\(j\) as a linear model \(f\) of the following covariates
\(\vec{x}_{ij}\)

@@

Using these posterior means, we then use the posterior predictive
distribution to obtain fitted/predicted values \(\widehat{y}\) of the
DBH for each focal tree using the \texttt{predict\_bayesian\_model()}.
These \(\widehat{y}\) can then be compared to the observed \(y\) DBH's
to compute the root mean-square error, a measure of a model's predictive
error which has the same units as the observed data \(y\).

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

@@

@@

@@

@@

@@

@@

@@

\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}
